---
title: "CITS4009 - Project 1"
author: Aaron Tan (23070356)
output: html_document
date: "2023-09-01"
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction 👋
The data set analyzed can be obtained from the Kaggle platform. It’s part of the “Global YouTube Statistics 2023" https://www.kaggle.com/datasets/nelgiriyewithana/global-youtube-statistics-2023

A collection of YouTube giants, this dataset offers a perfect avenue to analyze and gain valuable insights from the luminaries of the platform. With comprehensive details on top creators' subscriber counts, video views, upload frequency, country of origin, earnings, and more, this treasure trove of information is a must-explore for aspiring content creators, data enthusiasts, and anyone intrigued by the ever-evolving online content landscape. Immerse yourself in the world of YouTube success and unlock a wealth of knowledge with this extraordinary dataset.

Video link: https://youtu.be/PytUn6ANdnk

<br>
Table of each stage:

* [Basic Statistics](#basic-stats)
* [Data Cleaning/Transformation Stage](#data-clean)
* [Visualisations and Insights Stage](#visual-insights)
* [Business Queries](#bqueries)
* [Conclusion](#conclusion)

### Data loading, overview and set up: 
```{r}
library(shiny)
library(ggplot2)
library(gridExtra)
library(dplyr)
library(ca)
library(magrittr)
library(tidyverse)
library(data.table)
library(shinyWidgets)
library(leaflet)
```

#### Basic Statistics: {#basic-stats}
In this section, we demonstrate the iterative nature of EDA
by revisiting and refining our analysis at different stages.

Initial exploration:
<br>
This will perform a basic exploration of the data to get a sense of its structure.
We show the structure, summary, and head of the data to understand its basic characteristics.
```{r}
# Define a variable for the base directory
base_dir <- "D:/UWA/Data Science/Semester 2/CITS4009 - Computational Data Analysis/Assignment 1/"

# Define a variable for the file name
file_name <- "Global YouTube Statistics.csv"

# Combine the base directory and file name to create the full data path
data_path <- file.path(base_dir, file_name)
df <- fread(data_path, sep=",", encoding = "UTF-8")
```

```{r}
ui <- fluidPage(
  titlePanel("Exploration of Data Results:"),
  
  div(
    h4("Summary of results:"),
    p(HTML("There are <b>995 obs</b> with <b>28 variables.</b>")),
    p("There are 4 integers, 7 characters and 17 numeric columns (variables)."),
    p("Some of the data is not consistent like 0 video views, 'nan' and 'N/A'. I will data clean these values (Transform stage) before using a new dataset to properly do the EDA analysis."),
    
    class = "description-panel",
    p("Use the tabs below to view the structure (str), summary and head (first 10 datapoints per column) statistics."),
    hr()  
  ),
  
  mainPanel(
    tabsetPanel(
      tabPanel("str", verbatimTextOutput("str_output")),
      tabPanel("summary", verbatimTextOutput("summary_output")),
      tabPanel("head", tableOutput("head_output")),
    )
  )
)

# Display the 3 structures of data to understand its basic charactertistics
server <- function(input, output) {
  output$str_output <- renderPrint({
    str(df)
  })
  
  output$summary_output <- renderPrint({
    summary(df)
  })
  
  output$head_output <- renderTable({
    head(df, 10)
  })
}

shinyApp(ui, server)
```
### Data Cleaning/Transformation Stage 🧹 {#data-clean}
An important part of the ETL (Transform) Stage process before we do any analysis since this dataset does not contain perfect structured data
required for the analysis. This Stage is modularised into 5 processes:

* Columns underscore syntax should be consistent
* Replace all the zero, NA, and NaN values with their median
* Specifically get columns required 
* Remove outliers
* Setup any new dfs

#### Process 1: Columns underscore syntax should be consistent
Replace all the columns headings that are spaced to have replaced with _ and 
to make the column names all lowercases.
```{r}
clean_df <- df %>%
  rename_all(~ gsub(" ", "_", tolower(.)))
```

<br>

#### Process 2: Replace all the zero, NA, and NaN values with their median
Any missing values that are identified these 3 values are replaced each row based on its own column to its median.
```{r}
selected_columns <- c("video_views", "uploads", "country_rank", "gross_tertiary_education_enrollment_(%)", "population", "subscribers_for_last_30_days", "unemployment_rate", "channel_type_rank", "video_views_for_the_last_30_days", "urban_population", "created_year", "country")

# Calculate the median value for selected numeric columns and replace missing/zero values
for (col_name in selected_columns) {
  if (is.numeric(clean_df[[col_name]])) {
    clean_df[[col_name]][is.na(clean_df[[col_name]]) | is.nan(clean_df[[col_name]]) | clean_df[[col_name]] == 0] <- median(clean_df[[col_name]], na.rm = TRUE)
  }
}

```

<br>

#### Process 3: Specifically get columns required for a dataframe
Get the appropriate columns from clean_df
```{r}
select_df <- clean_df %>%
  select(subscribers, video_views, uploads, created_year, 
         category, lowest_monthly_earnings, highest_monthly_earnings,
         lowest_yearly_earnings, highest_yearly_earnings, population,
         unemployment_rate, urban_population)
```

<br>

#### Process 4: Remove outliers

```{r}
remove_outliers <- function(column) {
  q1 <- quantile(column, 0.25)
  q3 <- quantile(column, 0.75)
  iqr <- q3 - q1
  lower_bound <- q1 - 1.5 * iqr
  upper_bound <- q3 + 1.5 * iqr
  column[column < lower_bound | column > upper_bound] <- NA
  column
}

new_df <- select_df  %>%
  mutate(created_year = remove_outliers(created_year),
         subscribers = remove_outliers(subscribers),
         video_views = remove_outliers(video_views),
         uploads = remove_outliers(uploads),
         lowest_monthly_earnings = remove_outliers(lowest_monthly_earnings),
         highest_monthly_earnings = remove_outliers(highest_monthly_earnings),
         lowest_yearly_earnings = remove_outliers(lowest_yearly_earnings),
         highest_yearly_earnings = remove_outliers(highest_yearly_earnings)
         )
```

<br>

#### Process 5: Setup any new dfs

```{r}
bar_df <- clean_df %>%
  select("subscribers", "highest_yearly_earnings", "uploads", "unemployment_rate", "population", "youtuber", "country")

map_df <- clean_df %>%
  select("subscribers", "video_views", "uploads", "country","latitude","longitude")
```

### Visualisations and Insights Stage 🔍 {#visual-insights}
Now that the new dataset has been cleaned and transformed (ETL stage complete), we can properly perform Visualisations and Insights to interpret the following findings:
```{r}
ui <- fluidPage(
  titlePanel("Interactive Visualisations"),
  div(
     p("Select the following options: ")
  ),
  
  # Selection of plot types and perform if statements to determine whether it requires both x_attr and y_attr as 2 inpuits to appear
  sidebarLayout(
    sidebarPanel(
      selectInput("plot_type", "Select plot type:", choices = c("Scatter", "Map", "Dot Plot", "Bar Chart", "Pie Chart")),
      conditionalPanel(
        condition = "input.plot_type == 'Scatter'",
        selectInput("x_attr", "Select x-axis attribute:", names(new_df)),
        selectInput("y_attr", "Select y-axis attribute:", names(new_df))
      ),
      
      conditionalPanel(
        condition = "input.plot_type == 'Dot Plot' || input.plot_type == 'Bar Chart' || input.plot_type == 'Pie Chart'",
        selectInput("x_attr", "Select x-axis attribute:", names(new_df))
      ),
      
      conditionalPanel(
        condition = "input.plot_type == 'Map'",
        selectInput("d_attr", "Select size: ", names(new_df))
      )
    ),
  
   # Different plot output because not one plot is the same render function. Each plot output has its own render function.
   mainPanel(
      conditionalPanel(
        condition = "input.plot_type != 'Map'",
        plotOutput("plot", height = "400px", width = "600px")
      ),
      
      conditionalPanel(
        condition = "input.plot_type == 'Map'",
        leafletOutput("map", height = "400px", width = "600px")
      )
    )
  )
)

# Define the server
server <- function(input, output) {
  
  # If the user selects Map, it uses the renderLeaftlet function to render Leaflet package dependency.
  output$map <- renderLeaflet({
    d_attr <- input$d_attr
    lat <- map_df$latitude
    long <- map_df$longitude
    
    # Modifiy its radius_multiper based on its results since some d_attr has large values that needs to be downgraded
    if (d_attr == "subscribers") {
      radius_multiplier <- 1 / 8000000
      radius_color <- "darkred"
    } else if (d_attr == "video_views") {
      radius_multiplier <- 1 / 6000000000
      radius_color <- "darkblue"
    } else if (d_attr == "uploads"){
      radius_multiplier <- 1 / 10000
      radius_color <- "green"
    } else if (d_attr == "lowest_monthly_earnings"){
      radius_multiplier <- 1 / 10000
      radius_color <- "green"
    }
    
    # Display the leaflet map based on these values.
    leaflet() %>%
      addTiles() %>%
      setView(lng = mean(long, na.rm = TRUE), lat = mean(lat, na.rm = TRUE), zoom = 2) %>%
      addCircleMarkers(
        data = map_df,
        lng = long,
        lat = lat,
        radius = map_df[[d_attr]] * radius_multiplier,
        color = radius_color,
        fillOpacity = 0.7,
        popup = paste("Country: ", map_df$country, "<br>", d_attr, map_df[[d_attr]]) 
      )
  })
  
  # If the plot is not a map, it goes through the most common render type: renderPlot.
  output$plot <- renderPlot({
    if (input$plot_type == "Scatter") {
      x_attr <- input$x_attr
      y_attr <- input$y_attr
      
      ggplot(new_df, aes_string(x = x_attr, y = y_attr)) +
        geom_point() +
        geom_smooth(span = 0.1) +
        labs(title = paste(x_attr, " amount by ", y_attr))
    
    }else if(input$plot_type == "Dot Plot"){
      x_attr <- input$x_attr
      
      top_countries <- bar_df %>%
        group_by(country) %>%
        mutate(total_value = round(sum(bar_df[[x_attr]])/1000000)) %>%
        head(5)
        
      
      ggplot(top_countries, aes(x = total_value, y = country)) +
      geom_segment(aes_string(xend = 0, yend = "country"), color = "grey") +  # Add connecting lines
      geom_point(color = "dodgerblue", size = 4) +
      geom_text(aes(label = total_value), vjust = 3
                -2.5, hjust = -0.8, color = "black", size = 3) + 
      labs(title = "Top 5 Countries by sjns",
           x = "Subscribers",
           y = "Country") +
      theme_minimal() +
      scale_x_continuous(position = "top")
    
    }else if(input$plot_type == "Bar Chart"){
      x_attr <- input$x_attr
      y_attr <- bar_df$youtuber
      
      top_data <- head(bar_df[order(bar_df[[x_attr]], decreasing = TRUE), ], 10)
      top_data$x_values <- top_data[[x_attr]]
      
      top_data$y_attr_clean <- gsub("[^a-zA-Z0-9\\s.,!?&\'-]", " ", top_data$youtuber)
      top_data <- top_data[order(-top_data[[x_attr]]), ]
      
      ggplot(top_data, aes(x = x_values, y = y_attr_clean)) +
        geom_bar(stat = "identity", fill = "steelblue") +
        geom_text(aes(label = x_values), vjust = 0.5, hjust = 1.1, color = "white", size = 3.5, fontface = "bold") +
        labs(title = paste("Top 10 YouTube Channels by No.", x_attr), x = x_attr, y = "YouTuber") +
        theme_minimal()

      
    }else if(input$plot_type == "Pie Chart"){
      x_attr <- input$x_attr
  
      top_data <- head(bar_df[order(bar_df[[x_attr]], decreasing = TRUE), ], 10)
      top_data$x_values <- top_data[[x_attr]]
      
      top_data$y_attr_clean <- gsub("[^a-zA-Z0-9\\s.,!?&\'-]", " ", top_data$youtuber)
      top_data <- top_data[order(-top_data[[x_attr]]), ]
      
      ggplot(top_data, aes(x = "", y = x_values, fill = y_attr_clean)) +
        geom_bar(stat = "identity") +
        coord_polar(theta = "y") +
        labs(title = paste("Countries %", x_attr), x = NULL, y = NULL) +
        theme_minimal() +
        theme(legend.position = "bottom") 
    }
  })
}

# Run the Shiny app
shinyApp(ui, server)
```


### Business Queries: {#bqueries}
#### 1. Correlation between Number of Subscribers and Earnings (per year)
The correlation for both subscribers and highest_yearly_earnings showcase that vast majority of subscribers will earn
lower compared to those who have higher subscribers but that does not necessary mean it will generally have higher earnings.
For instance, some subscribers have low amount of subscribers but yet they achieved great earning results and some do have higher subscribers
but the earnings are low. Not much youtubers who can achieve a high amount of subscribers is necessary to achieve high earnings based on subscribers alone, it can affect the amount of earnings such as marketing campaigns to buy more products etc.

```{r, warning = FALSE }
ggplot(data = new_df, mapping = aes(x = new_df$subscribers, y = new_df$highest_yearly_earnings)) +
  geom_point() + 
  geom_smooth(method = "lm", formula = y ~ x) +
  labs(x = "No. of Subscribers", y = "$ per year") +
  ggtitle("Subscribers Earnings per Year")
```

#### 2. Top 10 YouTube Channels by No. Subscribers
T-Series has the highest amount of subscribers, followed by YouTube Movies and after that MrBeast. The next graph showcase whether those youtubers has the highest amount of earnings because of the number of subscribers that have?

```{r, warning = FALSE }
top_10_subscribers <- head(bar_df[order(bar_df$subscribers, decreasing = TRUE), ], 10)

top_10_subscribers$clean_youtuber <- gsub("[^a-zA-Z0-9\\s.,!?&\'-]", " ", top_10_subscribers$youtuber)

ggplot(top_10_subscribers, aes_string(x = "subscribers", y = "clean_youtuber")) +
        geom_bar(stat = "identity", fill = "steelblue") +
        geom_text(aes_string(label = "subscribers"), vjust = 0.5, hjust = 1.1, color = "white", size = 3.5, fontface = "bold") +
        labs(title = "Top 10 YouTubers by No. Subscribers", x = "Subscribers", y = "YouTuber")
```

#### 3. Top 10 YouTube Channels Max Year Earnings 
It not necessary true that the higher the amount of subscribers, the higher the earning potential that a Youtuber can produce. As you can see below, KIMPRO has the highest amount of earnings per year followed by DaFuq!?Boom! and after that T-Series. T-Series does showcase the highest amount of subscribers but it is the 3rd highest earning per year.
```{r}
top_10_earnings <- head(bar_df[order(bar_df$highest_yearly_earnings, decreasing = TRUE), ], 10)

top_10_earnings$clean_youtuber <- gsub("[^a-zA-Z0-9\\s.,!?&\'-]", " ", top_10_earnings$youtuber)

ggplot(top_10_earnings, aes_string(x = "highest_yearly_earnings", y = "clean_youtuber")) +
        geom_bar(stat = "identity", fill = "steelblue") +
        geom_text(aes_string(label = "highest_yearly_earnings"), vjust = 0.5, hjust = 1.1, color = "white", size = 3.5, fontface = "bold") +
        labs(title = "Top 10 YouTubers by their max earnings per Year ($)", x = "Max Earnings per Year ($)", y = "YouTuber")
```

#### 4. Top 10 Countries shows subscribers and max earnings comparsion
As you can see, every youtuber based on their country is added up their earnings and subscribers to represent which country contains the highest earnings and subscribers. We can definitely correlate that United States of America tends to have the highest earnings and subscribers, followed by India. Since majority of the world population's main language is English, it makes sense to see that majority of the population will be marketed towards English speakers.
```{r}
top10_country_max <- bar_df %>%
  group_by(country) %>%
  summarise(max_earnings = round(sum(highest_yearly_earnings)/10000000)) %>%
  arrange(desc(max_earnings)) %>%
  head(10)

top10_country_sub <- bar_df %>%
  group_by(country) %>%
  summarise(subscribers = round(sum(subscribers)/10000000)) %>%
  arrange(desc(subscribers)) %>%
  head(10)

p1 <- ggplot(top10_country_max, aes(x = max_earnings, y = country)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes_string(label = "max_earnings"), vjust = 0.5, hjust = 1.1,   color = "white", size = 3.5, fontface = "bold") +
  labs(title = "Top 10 Countries by Maximum Earnings",
       x = "Earnings (10M)",
       y = "Country")

p2 <- ggplot(top10_country_sub, aes(x = subscribers, y = country)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes_string(label = "subscribers"), vjust = 0.5, hjust = 1.1,   color = "white", size = 3.5, fontface = "bold") +
  labs(title = "Top 10 Countries by Subscribers",
       x = "Subscribers (10M)",
       y = "Country")

grid.arrange(p1, p2)
```

### Conclusion/Summary 📋 {#conclusion}
In Conclusion, United States of America and India seem to showcase exceptional results for having reputable Youtubers and the correlation between the earnings and subscribers shows little evidence to proof that if subscribers are higher then generally the earnings will be higher as well because the number of subscribers is not the only main factor that will determine it overall success (earnings). The top 10 YouTubers also clearly shows that even the highest amount of subscribers does not correlate to higher earnings. 

#### Improvements to be made
Bar Chart/Scatter Plot:

* Values can be displayed to cut into smaller digits (e.g. 1M) if the values gets too large, sort the YouTuber's data from biggest bar to lowest bar to see better comparison. 
* Ability to add a slider option not just fixed to only 10 YouTubers, it can showcase based on their selection for only the max. E.g I want to show 5 YouTubers.

The results only showcase 3 main variables:

* Subscribers
* Earnings
* Country


Dot-Plot and Pie Chart should be functional in the shiny-app interaction. Bar-Chart functionality is not completed as if I changed the x-axis value instead of subscribers to video_views, it should automatically renders its data properly.
To further improve it analysis, additional plot-types and enable functionality in shiny-app interaction and added additional data analysis for better insights to answer impactful queries. E.g. which category seem to have the most popular



