---
title: "CITS4009 - Project 1"
author: Aaron Tan (23070356)
output: html_document
date: "2023-08-04"
runtime: shiny
---

### Introduction 👋
The data set analyzed can be obtained from the Kaggle platform. It’s part of the “Global YouTube Statistics 2023"

A collection of YouTube giants, this dataset offers a perfect avenue to analyze and gain valuable insights from the luminaries of the platform. With comprehensive details on top creators' subscriber counts, video views, upload frequency, country of origin, earnings, and more, this treasure trove of information is a must-explore for aspiring content creators, data enthusiasts, and anyone intrigued by the ever-evolving online content landscape. Immerse yourself in the world of YouTube success and unlock a wealth of knowledge with this extraordinary dataset.

### Data loading, overview and set up: 
```{r}
library(shiny)
library(ggplot2)
library(gridExtra)
# library(ggthemes)
# library(numform)
#library(treemapify)
# library(timeDate)
# library(lubridate)
library(dplyr)
# library(reshape2)
library(ca)
library(magrittr)
library(tidyverse)
library(data.table)
library(shinyWidgets)
library(leaflet)
```


Load the main data:

```{r}
# data_path <- "D:/UWA/Data Science/Semester 2/CITS4009 - Computational Data Analysis/Assignment 1/Global YouTube Statistics.csv"
data_path <- "C:/Users/tanaaro/Desktop/youtube/Global YouTube Statistics.csv"
# df <- read.csv(data_path, sep=",", encoding = "UTF-8")
df <- fread(data_path, sep=",", encoding = "UTF-8")
```

```{r}
ui <- fluidPage(
  titlePanel("Exploration of Data Results:"),
  
  # Description panel using HTML elements
  div(
    h4("Summary of results:"),
    p(HTML("There are <b>995 obs</b> with <b>28 variables.</b>")),
    p("There are 4 integers, 7 characters and 17 numeric columns (variables)."),
    p("Some of the data is not consistent like 0 video views, 'nan' and 'N/A'. I will data clean these values (Transform stage) before using a new dataset to properly do the EDA analysis."),
    
    class = "description-panel",
    p("Use the tabs below to view the structure (str), summary and head (first 6 datapoints per column) statistics."),
    hr()  
  ),
  
  mainPanel(
    tabsetPanel(
      tabPanel("str", verbatimTextOutput("str_output")),
      tabPanel("summary", verbatimTextOutput("summary_output")),
      tabPanel("head", tableOutput("head_output")),
    )
  )
)

server <- function(input, output) {
  output$str_output <- renderPrint({
    str(df)
  })
  
  output$summary_output <- renderPrint({
    summary(df)
  })
  
  output$head_output <- renderTable({
    head(df, 10)
  })
}

shinyApp(ui, server)
```

### Data Cleaning/Transformation Stage 🧹
```{r}
# Stage 1: Columns underscore syntax should be consistent
clean_df <- df %>%
  rename_all(~ gsub(" ", "_", tolower(.)))

# Stage 2: Replace all the zero, NA, and NaN values with their median
selected_columns <- c("video_views", "uploads", "country_rank", "gross_tertiary_education_enrollment_(%)", "population", "subscribers_for_last_30_days", "unemployment_rate", "channel_type_rank", "video_views_for_the_last_30_days", "urban_population", "created_year", "country")

for (col_name in selected_columns) {
  if (is.numeric(clean_df[[col_name]])) {
    median_value <- median(clean_df[[col_name]][!is.na(clean_df[[col_name]]) & !is.nan(clean_df[[col_name]]) & clean_df[[col_name]] != 0])
    
    clean_df[[col_name]][is.na(clean_df[[col_name]]) | is.nan(clean_df[[col_name]]) | clean_df[[col_name]] == 0] <- median_value
  }
}
# Figure it out how to change decimal to integer for uploads, channel_type_rank column

# Stage 3: Specifically get columns required for scatterplot
scatter_df <- clean_df %>%
  # select(subscribers, video_views, uploads, country, created_year)
  select(subscribers, video_views, uploads, created_year)

# Stage 4: Remove outliers by using this function forumla:
remove_outliers <- function(column) {
  q1 <- quantile(column, 0.25)
  q3 <- quantile(column, 0.75)
  iqr <- q3 - q1
  lower_bound <- q1 - 1.5 * iqr
  upper_bound <- q3 + 1.5 * iqr
  column[column < lower_bound | column > upper_bound] <- NA
  column
}

# New df after removing outliers
scatter_df2 <- scatter_df %>%
  mutate(created_year = remove_outliers(created_year),
         subscribers = remove_outliers(subscribers),
         video_views = remove_outliers(video_views),
         uploads = remove_outliers(uploads))

# Stage 5: Setup any new dfs
bar_df <- clean_df %>%
  select("subscribers", "uploads", "country")

map_df <- clean_df %>%
  select("subscribers", "uploads", "country","latitude","longitude")
```


### EDA Analysis Stage 🔍
Now that the new dataset has been cleaned and transformed (ETL stage complete), we can properly perform EDA analysis to interpret the following findings:

```{r}
ui <- fluidPage(
  titlePanel("Interactive Visualisations 🌠"),
  div(
     p("Select the following options: ")
  ),
  
  sidebarLayout(
    sidebarPanel(
      selectInput("plot_type", "Select plot type:", choices = c("Scatter", "Map", "Bar Chart", "geom_boxplot", "geom_line")),
      conditionalPanel(
        condition = "input.plot_type == 'Scatter'",
        
        # Later figure out how to specifically outline numerically columns
        selectInput("x_attr", "Select x-axis attribute:", names(scatter_df2)),
        selectInput("y_attr", "Select y-axis attribute:", names(scatter_df2))
      ),
      conditionalPanel(
        condition = "input.plot_type == 'Bar Chart'",
        selectInput("x_attr", "Select x-axis attribute:", choices = c("subscribers", "uploads"))
      ),
      conditionalPanel(
        condition = "input.plot_type == 'Map'",
        selectInput("d_attr", "Select size: ", names(scatter_df2))
      )
    ),
    
   mainPanel(
      conditionalPanel(
        condition = "input.plot_type != 'Map'",
        plotOutput("plot", height = "400px", width = "600px")
      ),
      conditionalPanel(
        condition = "input.plot_type == 'Map'",
        leafletOutput("map", height = "400px", width = "600px")
      )
    )
  )
)

# Define the server
server <- function(input, output) {
  output$plot <- renderPlot({
    if (input$plot_type == "Scatter") {
      x_attr <- input$x_attr
      y_attr <- input$y_attr
      
      # print(paste("datatype: ", class(x_attr)))
      
      ggplot(scatter_df2, aes_string(x = x_attr, y = y_attr)) +
        geom_point() +
        geom_smooth(span = 0.1) +
        labs(title = paste(x_attr, " amount by ", y_attr))
    
    # Work on geom_bar
    }else if(input$plot_type == "Bar Chart"){
      x_attr <- input$x_attr
      values <- clean_df[[x_attr]]
      y_attr <- bar_df$country
      

      # y-axis for countries
      # x-axis for numeric data
      ggplot(bar_df, mapping = aes(x = values, y = y_attr)) +
        geom_bar(stat="identity", fill="gray") +
        geom_text(mapping=aes(label = x_attr), position = position_stack(vjust = 0.5), size=4) +
        ggtitle("Title here") +
        coord_flip() +
        labs(title = paste(x_attr, " amount by ", y_attr))

    # # Work on map last
    } else if (input$plot_type == "Map") {
      # d_attr <- as.numeric(input$d_attr)
      d_attr <- input$d_attr
      
      # summary(d_attr)
      # ("country","latitude","longitude", "uploads")
      
      # Identify numeric columns
      print(paste("d_attr: ", d_attr))
      # print(paste("map_df$d_attr: ", head(map_df[[d_attr]]), 10))
      
      print(paste("datatype: ", class(d_attr)))
      print(paste("datatype: ", class(map_df[[d_attr]])))
      
      map_df_numeric <- as.numeric(map_df[[d_attr]])
      print(paste("map_df_numeric ", class(map_df_numeric)))
      
      # print(paste("map_df$d_attr: ", map_df$d_attr))
      
      pal <- colorNumeric(
        palette = "Blues",
        domain = map_df_numeric
      )
      
      qpal <- colorQuantile("Blues", map_df_numeric, n = 7)
    
      leaflet() %>%
        addTiles() %>%
        addPolygons(
          data = d_attr,
          stroke = FALSE,
          smoothFactor = 0.2,
          fillOpacity = 1,
          fillColor = ~pal(map_df_numeric),
          popup = ~paste("<strong>Country:</strong><br>",
                         "<strong>Data:</strong>", map_df_numeric)
        )
    }
  })
}

# Run the Shiny app
shinyApp(ui, server)
```


### Business Queries:
#### 1. 2010 - 2015 between subscribers and video_views
The correlation for both subscribers and views showcase that the number of youtubers are primarily under 25% during the era in 2010-2015 making


```{r, warning=FALSE}
# Figure out how to change the ggplot size later
p1 <- ggplot(data = scatter_df2, mapping = aes(x = scatter_df2$subscribers, y = scatter_df2$created_year, color = (scatter_df2$created_year >= 2010 & scatter_df2$created_year <= 2015))) +
  geom_point() +
  scale_color_manual(values = c("FALSE" = "gray", "TRUE" = "red"),
  name = "Legend:",
  labels = c("Before 2010", "2010-2015")) +
  geom_smooth(span=0.1) +
  labs(x = "Subscribers Scatterplot", y = "Year") +
  ggtitle("Subscribers")

p2 <- ggplot(data = scatter_df2, mapping = aes(x = scatter_df2$video_views, y = scatter_df2$created_year, color = (scatter_df2$created_year >= 2010 & scatter_df2$created_year <= 2015))) +
  geom_point() +
  scale_color_manual(values = c("FALSE" = "gray", "TRUE" = "red"),
  name = "Legend:",
  labels = c("Before 2010", "2010-2015")) +
  geom_smooth(span=0.1) +
  labs(x = "Views", y = "Year") +
  ggtitle("Video_views")

grid.arrange(p1, p2, ncol=1)

```

#### 2. AI
As you can see....

#### 3. Softwware Engineering
As you can see....

#### 4.Cyber Security
As you can see....

#### 5. IT
As you can see....

### Conclusion/Summary 📋
In Conclusion, blah blah blah here


